The Werewolf Among Us: Humans vs LLMs in Multi-Agent Games

### 2025-05-06 _tex/paper.tex
Built site for gh-pages

  \begin{itemize}
  \tightlist
  \item
    \emph{Werewolf Among Us} (Lai et al., 2022), a collection of real
    human gameplay annotated with persuasion strategies,
  \item
    \emph{Werewolf Arena} (Bailis et al., 2024), a simulated environment
    where LLM agents play the game autonomously.
  \item
    While both studies show Werewolf generates complex strategic
    language, neither compares human and LLM behavior directly.
  \end{itemize}
\item
  Our project addresses this gap. We analyzed transcripts from both
  datasets, matched them by role, round, and persuasion strategy, and
  compared how humans and LLMs lie, persuade, and detect deception.
\item
  By annotating utterances with the same set of persuasive strategies,
  we clearly show how synthetic agents differ from or resemble humans
  when navigating deception in adversarial group interactions.
\end{itemize}

@@

We used the \emph{Werewolf Among Us} dataset (Lai et al., 2022), which
contains annotated dialogues from over 150 real games of One Night
Werewolf and Avalon. These games differ from classic Werewolf because
they have only one round of discussion and voting, do not eliminate
players during the game, and include many specialized roles beyond
Villager and Werewolf. The dataset provides detailed annotations of
persuasion strategies for each utterance, such as accusations, defenses,
and identity claims. We specifically used the textual transcriptions and
their strategy annotations for our comparisons.

@@

The \emph{Werewolf Arena} dataset (Bailis et al., 2024) features
simulated classic Werewolf games played by LLM agents without human
intervention. Unlike the one-round human games, these simulations
involve multiple rounds alternating between night (secret actions) and
day (open discussion). Each LLM agent is assigned a role (Villager,
Werewolf, Seer, Doctor) and receives context-specific prompts that guide
their strategic interactions.

>>>>

  \begin{itemize}
  \tightlist
  \item
    \emph{Werewolf Among Us}
    (\citeproc{ref-laiWerewolfUsMultimodal2022}{Lai et al. 2022}), a
    collection of real human gameplay annotated with persuasion
    strategies,
  \item
    \emph{Werewolf Arena}
    (\citeproc{ref-bailisWerewolfArenaCase2024}{Bailis, Friedhoff, and
    Chen 2024}), a simulated environment where LLM agents play the game
    autonomously.
  \item
    While both studies show Werewolf generates complex strategic
    language, neither compares human and LLM behavior directly.
  \end{itemize}
\item
  Our project addresses this gap. We analyzed transcripts from both
  datasets, matched them by role, round, and persuasion strategy, and
  compared how humans and LLMs lie, persuade, and detect deception.
\item
  By annotating utterances with the same set of persuasive strategies,
  we clearly show how synthetic agents differ from or resemble humans
  when navigating deception in adversarial group interactions.
\end{itemize}

@@

We used the \emph{Werewolf Among Us} dataset
(\citeproc{ref-laiWerewolfUsMultimodal2022}{Lai et al. 2022}), which
contains annotated dialogues from over 150 real games of One Night
Werewolf and Avalon. These games differ from classic Werewolf because
they have only one round of discussion and voting, do not eliminate
players during the game, and include many specialized roles beyond
Villager and Werewolf. The dataset provides detailed annotations of
persuasion strategies for each utterance, such as accusations, defenses,
and identity claims. We specifically used the textual transcriptions and
their strategy annotations for our comparisons.

@@

The \emph{Werewolf Arena} dataset
(\citeproc{ref-bailisWerewolfArenaCase2024}{Bailis, Friedhoff, and Chen
2024}) features simulated classic Werewolf games played by LLM agents
without human intervention. Unlike the one-round human games, these
simulations involve multiple rounds alternating between night (secret
actions) and day (open discussion). Each LLM agent is assigned a role
(Villager, Werewolf, Seer, Doctor) and receives context-specific prompts
that guide their strategic interactions.
